\Large \bf FlexAlloc: Dynamic Memory Partitioning\\ between the Corevisor and Hostvisor in SeKVM

### 2023-12-24 paper.tex
Added iterator diagrams and elaborations in section 2

@@

%for single author (just remove % characters)
\author{
{\rm Matthew Roger Nelson}\\
Columbia University
\and
{\rm Ryan Wee}\\
Columbia University
\and
{\rm Zeheng Yang}\\
Columbia University
} % end author

@@

@@

As described above, the original implementation of SeKVM statically
allocates sixteen contiguous 8M regions to the corevisor. When a VM is created,
the corevisor uses one of these contiguous 8M regions for the VM's stage-2
page table. To implement a system of dynamic allocation, we want the hostvisor
to dynamically reserve a contiguous 8M region at runtime when a VM is created, and then
pass this memory region to the corevisor. The kernel macro \textit{alloc\_pages}
is responsible for allocating a large number of contiguous physical pages~\cite{kerneldocs}.
Examining the implementation of this macro reveals that it eventually
calls \textit{\_\_alloc\_pages\_nodemask}, which checks that the order requested
is strictly less than \textit{MAX\_ORDER}~\cite{page_alloc.c}. \textit{MAX\_ORDER}
is in turn defined to be 11~\cite{mmzone.h}. This means that we cannot allocate
a contiguous region of memory larger than $2^{10} * 4K = 4M$. This
limit is set by the kernel because \textit{alloc\_pages} uses the buddy allocator,
which recursively splits available memory into half to find the best possible position for
a contiguous region of memory~\cite{kerneldocs}. However, this also makes it
undesirable to continuously allocate huge contiguous regions of memory.

@@

>>>>

\usepackage{graphicx}
\graphicspath{ {./imgs/} }

@@

%for single author (just remove % characters)
\author{
{\rm Matthew Nelson}\\
Columbia University
\and
{\rm Ryan Wee}\\
Columbia University
\and
{\rm Zeheng Yang}\\
Columbia University
} % end author

@@

Although the metadata structs have a much smaller memory footprint than the stage-2 page tables,
we designed and partially implemented a fully generalized dynamic allocation interface that
could be used to support dynamic allocation for all metadata structs which we call the
Dynamic Page Manager (DPM). DPM handles all of the work required to receive and return
memory from the hostvisor allowing other "services" in EL2 to exchange memory with DPM using
a very similiar interface to the alloc_pages and free_pages interface available in EL1. We describe
this interface in detail in Section 4.

@@

As described above, the original implementation of SeKVM statically
allocates sixteen contiguous 8M regions to the corevisor (Figure \ref{fig:sekvmvmregions}). When a VM is created,
the corevisor uses one of these contiguous 8M regions for the VM's stage-2
page table. To implement a system of dynamic allocation, we want the hostvisor
to dynamically reserve a contiguous 8M region at runtime when a VM is created, and then
pass this memory region to the corevisor. The kernel macro \textit{alloc\_pages}
is responsible for allocating a large number of contiguous physical pages~\cite{kerneldocs}.
Examining the implementation of this macro reveals that it eventually
calls \textit{\_\_alloc\_pages\_nodemask}, which checks that the order requested
is strictly less than \textit{MAX\_ORDER}~\cite{page_alloc.c}. \textit{MAX\_ORDER}
is in turn defined to be 11~\cite{mmzone.h}. This means that we cannot allocate
a contiguous region of memory larger than $2^{10} * 4K = 4M$. This
limit is set by the kernel because \textit{alloc\_pages} uses the buddy allocator,
which recursively splits available memory into half to find the best possible position for
a contiguous region of memory~\cite{kerneldocs}. However, this also makes it
undesirable to continuously allocate huge contiguous regions of memory.

\begin{figure}[h!tbp]
\caption{SeKVM VM Stage-2 Page Table Region Layout, Statically Allocated}
\label{fig:sekvmvmregions}
\centering
    \includegraphics{sekvm_vm_region}
\end{figure}

@@

\begin{figure}[h!tbp]
\caption{FlexAlloc: First 1MB Region Metadata Setup}
\label{fig:firstmbregion}
\centering
    \includegraphics{firstmbregion}
\end{figure}

\begin{figure}[h!tbp]
\caption{FlexAlloc: Second to Eigth 1MB Region Metadata Setup}
\label{fig:otherregions}
\centering
    \includegraphics{otherregions}
\end{figure}

Each level's iterators have a set number of regions from which to pull additional pages from. Lower
levels in the table have larger regions than higher levels in the table. When the current iterator
over a given region is exhausted, the iterator will jump to the next available region and pick up
where it left off.

\begin{verbatim}
static u64 inline get_pmd_next(u32 vmid) {
    ...
    // pool_index: available iterator
    // we are using
    pool_index =
      el2_data->vm_info[vmid].pte_pool_index;
    // pool_start: beginning of the region
    // the pool_index corresponds to
    pool_start =
      el2_data->vm_info[vmid].
      pte_pool_starts[pool_index];
    // used_pages: number of pages in this
    // region we have exhausted
    used_pages = el2_data->vm_info[vmid]
      .pte_used_pages_vm[pool_index];

    // If we've used up all the pages and we
    // have another region, then use the new region
    if (used_pages == PTE_ITER_MAX_PAGES &&
        pool_index + 1 < PTE_USED_ITER_COUNT) {
        pool_index += 1;
        el2_data->vm_info[vmid].
          pte_pool_index = pool_index;

        // Set the pool_start to the next available
        // PTE iterator
        pool_start = el2_data->vm_info[vmid].
          pte_pool_starts[pool_index];
        used_pages = el2_data->vm_info[vmid].
          pte_used_pages_vm[pool_index];
    }

    // Make sure we have not exhausted all pages
    // across all regions
    if (used_pages == PTE_ITER_MAX_PAGES &&
        pool_index == PTE_USED_ITER_COUNT - 1) {
        __hyp_panic();
    }

    // Return the physical address of the next
    // page that this iterator can allocate
    return pool_start +
      (used_pages * PAGE_SIZE);
}
\end{verbatim}
